Computing solution space properties by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-31 paper/paper.tex
Updates from Overleaf

This set algebra can serve as the coefficients in \Eq{eq:polynomial} to enumerate independent sets of all different sizes, \Eq{eq:countingtropical} to enumerate all MISs, or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
As long as the coefficients in a truncated polynomial are members of a commutative semiring, the polynomial itself is a commutative semiring.
For example, to enumerate only the MISs we can define a combined element type $s_{k}\infty^k$, where the coefficients follow the algebra in \Eq{eq:set} and the exponents follow the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\eqname{P1+SN}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm P1+SN}(\{\boldsymbol{e}_{i}\}\infty^1) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}\infty^1
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm P1+SN} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The enumeration of all MIS configurations corresponds to the contraction of this tensor network.
However, direct contraction might incur significant space overheads for keeping too many intermediate states irrelevant to the final maximum independent sets.
This issue can be avoided by using the bounding technique in \App{sec:bounding}.
One may also be interested in a more studied maximal independent sets~\cite{Bron1973, Eppstein2010, Johnson1988} enumeration. We will discuss this issue in \App{sec:maximal} since it requires using a different tensor network structure for the maximal independent set problem.

If one is interested in obtaining only a single MIS configuration, he can just keep one configuration in each tensor element to save the computational effort.
By replacing the sets of bit strings in \Eq{eq:set} with a single bit string, we have the following algebra

@@

Let us characterize the independent set problem on graph $G=(V, E)$ as an energy model
\begin{equation}\label{eq:eng}
    \mathcal{E}(G, s) = -\sum_{i\in V(G)} w_i s_i + \infty \sum_{\langle i,j\rangle \in E(G)}s_i s_j
\end{equation}
where $s_i$ is a spin on vertex $i \in V$ and $w_i$ is an onsite energy term associated with it.
The first term is the energy that corresponds to the inverse of the independent set size and the second term describes the independence contraint.
In physical systems, this contraint term usually corresponds to the Rydberg blockade~\cite{Pichler2018, Ebadi2022} in cold atom arrays or the repulsive force in hard core lattice model~\cite{Dyre2016, Fernandes2007}.
Its partition function is defined as
\begin{equation}\label{eq:partition}
    \begin{split}
    Z(G, \beta) = \sum_{s}e^{-\beta \mathcal{E}(G, s)} = \sum_{s\in \mathcal{I}(G)} e^{\beta \sum w_i s_i}\\
         = \sum_{k=0}^{\alpha(G)}a(k) e^{\beta k}  \qquad \quad (k = \sum w_i s_i)
    \end{split}
\end{equation}
where $\mathcal{I}(G)$ is the set of independent sets of graph $G$.
$\alpha(G)$ is the absolute value of the minimum energy (maximum independent set size).
$a(k)$ is the number of spin configurations with energy $-k$ (independent sets of size $k$).
The partition function can be evaluated as a tensor network contraction,
where the tensor network is constructed by placing a vertex tensor on each spin $i$
\begin{equation}
    W(e^{\beta w_i}) = \left(\begin{matrix}
        1 \\
        e^{\beta w_i}
    \end{matrix}\right),
\end{equation}
and an edge tensor on each bond
\begin{equation}
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right),
\end{equation}
where the $0$ in the bond tensor coming from $e^{-\beta\infty}$ in the second term of \Eq{eq:eng}.
By letting $x = e^{\beta}$, we get the tensor network for computing the independence polynomial as described by \Eq{eq:vertextensor} and \Eq{eq:edgetensor},
and by letting $w_i=1$, the second line of \Eq{eq:partition} is the independence polynomial.

@@

In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice, either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
Hence a tensor network can be represented as a open simple graph, where a tensor is mapped to a vertex, a label shared by two input tensors is mapped to a normal edge and a label appears in the output tensor is mapped to an open edge.
A standard tensor network notation is equivalent to the generalized tensor network by introducing $\delta$ tensors, where a $\delta$ tensor of rank $d$ is defined as
\begin{equation}
    \delta_{i_1, i_2,\ldots,i_d} = \begin{cases}
        1, & i_1=i_2=\ldots =i_d,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}

In the following, we are going to show introducing $\delta$ tensors might increase the contraction complexity of a tensor network.
Let us consider the following King's graph.

@@

By mapping the independent set problem to a standard tensor network with $\delta$ tensors,
we have the following graphical representation.

@@

In this diagram, the circle on each vertex in the original graph is a $\delta$ tensor of rank $8$ or less.
If we contract this tensor network in a naive column-wise order, the maximum intermediate tensor has rank $\sim3L$, requiring a storage of size $\approx 2^{3L}$.
If we relax the restriction that each label must appears exactly twice, we have the following hypergraph representation of a generalized tensor network.

\centerline{\begin{tikzpicture}
    \def\r{0.08}
    \def\a{0.07}
    \def\L{0.6}
    \def\l{0.1}
    \def\sql{0.24}
    \pgfmathsetseed{2}
    \foreach[evaluate={\cr=0.1+0.5*Mod(\x,2)}] \x in {1,...,5}
        \foreach[evaluate={\cg=0.1+0.3*Mod(\y,2); \cy=0.5-0.5*Mod(\y,2)}] \y in {1,...,5}{
            \edef\R{\pdfuniformdeviate 255}
            \edef\G{\pdfuniformdeviate 255}
            \edef\B{\pdfuniformdeviate 255}
            \xdefinecolor{MyColor}{RGB}{\R,\G,\B}
            \ifnum \x < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x+0.5,\y);
                \ifnum \y < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x+0.3,\y+0.3);
                \fi
                \ifnum \y > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x+0.3,\y-0.3);
                \fi
            \fi
            \ifnum \x > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x-0.5,\y);
                \ifnum \y < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x-0.7,\y+0.7);
                \fi
                \ifnum \y > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x-0.7,\y-0.7);
                \fi
            \fi
            \ifnum \y < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x,\y+0.5);
            \fi
            \ifnum \y > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x,\y-0.5);
            \fi
        }
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,5}{
            %\filldraw[fill=black] (\x,\y) circle [radius=0.7*\r];
        }
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,5}{
            \ifnum \y < 5
                \filldraw[fill=black] (\x,\y+0.5) circle [radius=\r];
                \filldraw[fill=black] (\y+0.5,\x) circle [radius=\r];
            \fi
        }
    \foreach \x in {1,...,4}
        \foreach \y in {1,...,4}{
            \filldraw[fill=black] (\x+0.3,\y+0.3) circle [radius=\r];
            \filldraw[fill=black] (\y+0.3,\x+0.7) circle [radius=\r];
        }
    \tikzset{decoration={snake,amplitude=.4mm,segment length=2mm,
                    post length=0mm,pre length=0mm}}
    \draw [decorate] (2.7, 0.5) -- (2.7, 5.5);
\end{tikzpicture}}
Here, we use different colors to distinguish different hyperedges.
A vertex tensor always has rank $1$ and is not shown here since it does not change the contraction complexity.
If we contract this tensor network in the column-wise order, the maximum intermediate tensor rank is $\sim L$, which can be seen by counting the number of colors at the cut.

\section{Hard problems and tensor networks}\label{sec:otherproblems}
\subsection{Maximal independent sets and maximal cliques}\label{sec:maximal}
Since finding maximal cliques of a graph is equivalent to finding the maximal independent sets of its complement graph, in the following,
we limit our discussion to finding maximal independent sets.
Let us denote the neighborhood of a vertex $v$ as $N(v)$ and the closed neighborhood of $v$ as $N[v] = N(v)\cup \{v\}$.
A maximal independent set $I_m$ is an independent set where there does exist a vertex $v \in V$ such that $I_m \cap N[v]  = \emptyset$,
i.e. an independent set that can not become a larger one by adding a new vertex.
To characterize the maximal independence restriction, one can quantify over a local set $N[v]$ with a local tensor:
\begin{equation}\label{eq:maximal}
    T(x_v, w_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
        s_vx_v^{w_v} & s_1=s_2=\ldots=s_{|N(v)|}=0,\\
        1-s_v& \text{otherwise}.\\
    \end{cases}
\end{equation}
Intuitively, it means if all the neighbourhood vertices are not in $I_{m}$, i.e., $ s_1=s_2=\ldots=s_{|N(v)|}=0$, then $v$ should be in $I_{m}$ and contribute a factor $x_{v}$,
otherwise, if any of the neighbourhood vertices is in $I_{m}$, then $v$ cannot be in $I_{m}$.
As an example, for a vertex of degree 2, the resulting rank-3 tensor is
\begin{equation}
    T(x_v, w_v)=\left(\begin{matrix}
    \left(\begin{matrix}
        ~~0 &~1 \\
        ~~1 &~1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v^{w_v} &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}

Let us consider the example in~\Sec{eg:tensorcontraction}: its corresponding tensor network structure for computing the maximal independent polynomial becomes

@@

One can see that the average rank of a tensor is increased.
The computational complexity of this new tensor network contraction is often greater than the one for computing the independence polynomial.

By contracting this tensor network with generic element types,
we can compute the maximal independent set properties such as the maximal independence polynomial and the enumeration of maximal independent sets.
Similar to the independence polynomial, the maximal independence polynomial counts the number of maximal independent sets of various sizes~\cite{Hu2017}.
%which can be helpful for understanding why the program for solving MIS can be trapped in a local minimum.
Let $G=(V,E)$ be a graph, its maximal independence polynomial is defined as
\begin{equation}
D_{m}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$.
Comparing with the independence polynomial in \Eq{eq:idpdef}, we have $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $D_m(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012, Manne2013},
which to our knowledge, the best algorithm gives a time complexity $O(1.3642^{|V|})$~\cite{Gaspers2012}.
If we want to find an MIS, $b_{k}$ can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\cite{Ebadi2022}.
%The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017}.

We show the benchmark of computing the maximal independent set properties in \Fig{fig:benchmark-maximal},
including a comparison to the Bron-Kerbosch algorithm from Julia package Graphs~\cite{Graphs}.
The treewidth of this tensor network is significantly larger, so benchmark with only a smaller graph size is feasible.
The time for the tensor network approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
Due to the memory limit, this Bron-Kerbosch algorithm is only feasible up to a graph size around $70$.

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig2.pdf}
    \caption{Benchmark results for computing different properties of maximal independent sets on a random three regular graph with different tensor element types.
    (a) treewidth versus the number of vertices for the benchmarked graphs. 
    (b) The computing time for calculating the number of independent sets and enumerate all MISs.
    }
    \label{fig:benchmark-maximal}
\end{figure}

\subsection{Matching problem}
A $k$-matching in a graph $G=(V,E)$ is a set of $k$ edges, no two of which have a vertex in common.
We map an edge $(u, v) \in E$ to a degree of freedom $\langle u, v\rangle \in \{0, 1\}$ in a tensor network,
where $1$ means this edge is in the set and $0$ otherwise.
To characterize the matching, one can define a tensor of rank $d(v) = |N(v)|$ on vertex $v$
\begin{equation}
    W_{\langle v, n_1\rangle, \langle v, n_2 \rangle, \ldots, \langle v, n_{d(v)}\rangle} = \begin{cases}
        1, & \sum_{i=1}^{d(v)} \langle v, n_i \rangle \leq 1,\\
        0, & \text{otherwise},
    \end{cases}
\end{equation}
and a tensor of rank $1$ on the bond
\begin{equation}
    B(w_{\langle u,v \rangle})_{\langle u, v\rangle} = \begin{cases}
    1, & \langle u, v \rangle = 0 \\
    x^{w_{\langle u,v \rangle}}_{\langle u, v\rangle}, & \langle u, v \rangle = 1,
\end{cases}
\end{equation}
where label $\langle v, u \rangle$ is equivalent to $\langle u,v\rangle$.
The vertex tensor specifies the constraint that a vertex cannot be shared by two edges in the matched set,
while an edge tensor carries the weights as the target to optimize.
Let $x_{\langle u,v\rangle}^{w_{\langle u,v\rangle}}=x$, the tensor network contraction gives us the matching polynomial
\begin{equation}
    M(G, x) = \sum\limits_{k=1}^{|V|/2} c_k x^k,
\end{equation}
where $k$ is the size of a matched edges set, and a coefficients $c_k$ is the number of $k$-matchings.

\subsection{Vertex coloring}
Let $G=(V,E)$ be a graph, a vertex coloring is an assignment of colors to each vertex $v\in V$ such that no edge connects two identically colored vertices. 
In a $k$-coloring problem, the number of different colors can be used is limited to less or equal to $k$.
Let us use the 3-coloring problem as an example to show how to characterize with local tensors.
For each vertex $v \in V$, we associate a degree of freedom $c_v\in\{0,1,2\}$ to it.
Then we define a vertex tensor labelled by it as
\begin{equation}
    W = \left(\begin{matrix}
        r_v\\
        g_v\\
        b_v
    \end{matrix}\right),
\end{equation}
where $r_v$, $g_v$ and $b_v$ are colors for labeling the configurations.
For each edge $(u, v)$, we specify the constraint carrying a weight $w_{uv}$ on it by defining an edge tensor labelled by $(c_u, c_v)$ as
\begin{equation}
    B(x^{w_{uv}}) = \left(\begin{matrix}
        0 & x^{w_{uv}} & x^{w_{uv}}\\
        x^{w_{uv}} & 0 & x^{w_{uv}}\\
        x^{w_{uv}} & x^{w_{uv}} & 0
    \end{matrix}\right).
\end{equation}
The number of possible colorings can be obtained by contracting this tensor network by setting $r_v, g_v$ and $b_v$ to $1$.
Let $x^{w_{uv}} = x$, we have a graph polynomial,
in which the $k$-th coefficient is the number of coloring with $k$ bonds satisfied.
If a graph is colorable, the maximum order of this polynomial should be equal to the number of edges in this graph.
Similarly, one can define the $k$-coloring problem on edges by defining the tensor network on the line graph of $G$.

\subsection{Cutting problem}
In graph theory, a cut is a partition of the vertices of a graph into two disjoint subsets,
which is also known as the boolean spin glass problem in statistic physics.
Let $G=(V,E)$ be a graph, to reduce the cutting problem on $G$ to the contraction of a tensor network, we first define a boolean degree of freedom $s_v\in\{0, 1\}$ for each vertex $v\in V$.
Then for each edge $(u,v)\in E$, we define an edge matrix labelled by $s_us_v$ as
\begin{equation}
    B(x, w_{uv}) = \left(\begin{matrix}
        1 & x_{v}^{w_{uv}}\\
        x_{u}^{w_{uv}} & 1
    \end{matrix}\right),
\end{equation}
where variables $x_u$ and $x_v$ represents a cut on this edge or a domain wall in an spin glass.
Let $x_u^{w_{uv}} = x_v^{w_{uv}} = x$, we have a graph polynomial similar to previous ones,
where its $k$th coefficient is two times the number of cuts with size $k$ (i.e. cutting $k$ edges).

>>>>

This set algebra can serve as the coefficients in \Eq{eq:polynomial} to enumerate independent sets of all different sizes, \Eq{eq:countingtropical} to enumerate all MISs, or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
As long as the coefficients in a truncated polynomial are members of a commutative semiring, the polynomial itself is a commutative semiring.
For example, to enumerate only the MISs we can define a combined element type $s_{k}\infty^k$, where the coefficients follow the algebra in \Eq{eq:set} and the exponents follow the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\eqname{P1+SN}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm P1+SN}(\{\boldsymbol{e}_{i}\}\infty^1) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}\infty^1
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm P1+SN} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The enumeration of all MIS configurations corresponds to the contraction of this tensor network.
However, direct contraction might have significant space overheads for keeping too many intermediate states irrelevant to the final maximum independent sets.
We introduce the bounding technique in \App{sec:bounding} to avoid this issue.
One may also be interested in a more studied maximal independent sets~\cite{Bron1973, Eppstein2010, Johnson1988} enumeration.
It will be discussed in \App{sec:maximal} since it requires using a different tensor network structure.

If one is interested in obtaining only one MIS configuration, he can just keep one configuration in each tensor element to save the computational effort.
By replacing the sets of bit strings in \Eq{eq:set} with a single bit string, we have the following algebra

@@

Let us characterize the independent set problem on graph $G=(V, E)$ as an energy model with two parts
\begin{equation}\label{eq:eng}
    \mathcal{E}(G, s) = -\sum_{i\in V} w_i s_i + \infty \sum_{\langle i,j\rangle \in E}s_i s_j
\end{equation}
where $s_i$ is a spin on vertex $i \in V$ and $w_i$ is an onsite energy term associated with it.
The first part corresponds to the negative independent set size and the second part describes the independence contraint, which corresponds to the Rydberg blockade~\cite{Pichler2018, Ebadi2022} in cold atom arrays or the repulsive force in hard core lattice model~\cite{Dyre2016, Fernandes2007}.
The partition function is defined as
\begin{equation}\label{eq:partition}
    \begin{split}
    Z(G, \beta) = \sum_{s}e^{-\beta \mathcal{E}(G, s)} = \sum_{s\in \mathcal{I}(G)} e^{\beta \sum w_i s_i}\\
         = \sum_{k=0}^{\alpha(G)}a(k) e^{\beta k}  \qquad \quad (k = \sum w_i s_i)
    \end{split}
\end{equation}
where $\mathcal{I}(G)$ is the set of independent sets of graph $G$, $\alpha(G)$ is the absolute value of the minimum energy (maximum independent set size), $a(k)$ is the number of spin configurations with energy $-k$ (independent sets of size $k$).
The partition function can be expressed as a tensor network by placing a vertex tensor on each spin $i$
\begin{equation}
    W(e^{\beta w_i}) = \left(\begin{matrix}
        1 \\
        e^{\beta w_i}
    \end{matrix}\right),
\end{equation}
and an edge tensor on each bond
\begin{equation}
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right),
\end{equation}
where the $0$ in the bond tensor coming from $e^{-\beta\infty}$ in the second term of \Eq{eq:eng}.
By letting $x = e^{\beta}$, we get the tensor network for computing the independence polynomial as described by \Eq{eq:vertextensor} and \Eq{eq:edgetensor}.
If we further let $w_i=1$, the second line of \Eq{eq:partition} is equivalent to the independence polynomial.

@@

In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice: either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
Hence a tensor network can be represented as a open simple graph, where an input tensor is mapped to a vertex, a label shared by two input tensors is mapped to an edge and a label appears in the output tensor is mapped to an open edge.
A standard tensor network notation is equivalent to the generalized tensor network in representation power.
A generalized tensor network can be converted to a standard one by adding a $\delta$ tensors at each hyperedge, where a $\delta$ tensor of rank $d$ is defined as
\begin{equation}
    \delta_{i_1, i_2,\ldots,i_d} = \begin{cases}
        1, & i_1=i_2=\ldots =i_d,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}

In the following example, we will show this conversion might increase the contraction complexity of a tensor network.
Let us consider the following King's graph.

@@

The generalized tensor network for solving the MIS problem on this graph has the following hypergraph representation, where we use different colors to distinguish different hyperedges.

\centerline{\begin{tikzpicture}
    \def\r{0.08}
    \def\a{0.07}
    \def\L{0.6}
    \def\l{0.1}
    \def\sql{0.24}
    \pgfmathsetseed{2}
    \foreach[evaluate={\cr=0.1+0.5*Mod(\x,2)}] \x in {1,...,5}
        \foreach[evaluate={\cg=0.1+0.3*Mod(\y,2); \cy=0.5-0.5*Mod(\y,2)}] \y in {1,...,5}{
            \edef\R{\pdfuniformdeviate 255}
            \edef\G{\pdfuniformdeviate 255}
            \edef\B{\pdfuniformdeviate 255}
            \xdefinecolor{MyColor}{RGB}{\R,\G,\B}
            \ifnum \x < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x+0.5,\y);
                \ifnum \y < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x+0.3,\y+0.3);
                \fi
                \ifnum \y > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x+0.3,\y-0.3);
                \fi
            \fi
            \ifnum \x > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x-0.5,\y);
                \ifnum \y < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x-0.7,\y+0.7);
                \fi
                \ifnum \y > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x-0.7,\y-0.7);
                \fi
            \fi
            \ifnum \y < 5
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x,\y+0.5);
            \fi
            \ifnum \y > 1
                \draw [thick, MyColor, opacity=1.0, line cap=round] (\x,\y) -- (\x,\y-0.5);
            \fi
        }
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,5}{
            %\filldraw[fill=black] (\x,\y) circle [radius=0.7*\r];
        }
    \foreach \x in {1,...,5}
        \foreach \y in {1,...,5}{
            \ifnum \y < 5
                \filldraw[fill=black] (\x,\y+0.5) circle [radius=\r];
                \filldraw[fill=black] (\y+0.5,\x) circle [radius=\r];
            \fi
        }
    \foreach \x in {1,...,4}
        \foreach \y in {1,...,4}{
            \filldraw[fill=black] (\x+0.3,\y+0.3) circle [radius=\r];
            \filldraw[fill=black] (\y+0.3,\x+0.7) circle [radius=\r];
        }
    \tikzset{decoration={snake,amplitude=.4mm,segment length=2mm,
                    post length=0mm,pre length=0mm}}
    \draw [decorate] (2.7, 0.5) -- (2.7, 5.5);
\end{tikzpicture}}
Here, vertex tensors are not shown here because they can be absorbed into an edge tensor hence do not change the contraction complexity.
If we contract this tensor network in the column-wise order, the maximum intermediate tensor has rank $\sim L$, which can be seen by counting the number of colors at the cut.

By adding $\delta$ tensors to hyperedges, we have the standard tensor network represented as the following simple graph.

@@

In this diagram,the additional $\delta$ tensors can have ranks upto $8$.
If we still contract this tensor network in a column-wise order, the maximum intermediate tensor has rank $\sim3L$, i.e. the space complexity is $\approx 2^{3L}$.

\section{Hard problems and tensor networks}\label{sec:otherproblems}
\subsection{Maximal independent sets and maximal cliques}\label{sec:maximal}
In this section, we focus the discussion on the maximal independent sets problem since finding maximal cliques of a graph is equivalent to finding the maximal independent sets of the complement of the graph.
Let $G=(V,E)$ be a graph, we denote the neighborhood of a vertex $v\in V$ as $N(v)$.
A maximal independent set $I_m$ is an independent set that there does exist a vertex $v \in V$ such that $I_m \cap (\{v\} \cup N[v])  = \emptyset$, i.e. an independent set that can not become a larger one by adding a new vertex.
To characterize the maximal independence restriction, we defined a tensor on each $\{v\}\cup N(v)$ as
\begin{equation}\label{eq:maximal}
    T(x_v, w_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
        s_vx_v^{w_v} & s_1=s_2=\ldots=s_{|N(v)|}=0,\\
        1-s_v& \text{otherwise}.\\
    \end{cases}
\end{equation}
It means if none of the neighbours of $v$ are in $I_{m}$ ($ s_1=s_2=\ldots=s_{|N(v)|}=0$), then $v$ must be in $I_{m}$ and contribute a factor $x_{v}$, otherwise, if any of the neighbourhood vertices is in $I_{m}$, then $v$ cannot be in $I_{m}$.
For a vertex $v$ having degree 2, this tensor has the following form
\begin{equation}
    T(x_v, w_v)=\left(\begin{matrix}
    \left(\begin{matrix}
        ~~0 &~1 \\
        ~~1 &~1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v^{w_v} &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}

Let us consider the graph in Example~\ref{eg:tensorcontraction}, the corresponding tensor network structure for computing the maximal independent polynomial has the following hypergraph representation.

@@

%The contraction complexity of this tensor network is often greater than the one for the independence set problem.
By contracting this tensor network with generic element types,
we can compute the maximal independent set properties such as the maximal independence polynomial and the enumeration of maximal independent sets.
The maximal independence polynomial is defined as
\begin{equation}
D_{m}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$.
Comparing with the independence polynomial in \Eq{eq:idpdef}, we have $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $D_m(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012, Manne2013},
which to our knowledge, the best algorithm gives a time complexity $O(1.3642^{|V|})$~\cite{Gaspers2012}.
%If we want to find an MIS, $b_{k}$ can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\cite{Ebadi2022}.
%The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017}.

We show the benchmark of computing the maximal independent set properties on $3$-regular graphs in \Fig{fig:benchmark-maximal},
including a comparison to the Bron-Kerbosch algorithm from Julia package \href{https://github.com/JuliaGraphs/Graphs.jl}{Graphs}~\cite{Graphs}.
\Fig{fig:benchmark-maximal} (a) shows the space and time complexities of tensor contraction, which are typically larger than those for the independence set problem.
In \Fig{fig:benchmark-maximal} (b), one can see counting maximal independent sets are much more efficient than enumerating them, while our generic tensor network approach runs slightly faster than the Bron-Kerbosch approach in enumerating all maximal independent sets.

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig2.pdf}
    \caption{Benchmarks of computing different solution space properties of the maximal independent sets (ISs) problem on random three regular graphs at different sizes.
    (a) time and space complexity of tensor network contraction. 
    (b) The wall clock time for counting and enumeration of maximal ISs. The missing data is due to due to the memory limit.
    }
    \label{fig:benchmark-maximal}
\end{figure}

\subsection{Matching problem}
A $k$-matching in a graph $G=(V,E)$ is a set of $k$ edges that no two of which have a vertex in common.
We map an edge $(u, v) \in E$ to a degree of freedom $\langle u, v\rangle \in \{0, 1\}$ in a tensor network, where $1$ means an edge is in the set and $0$ otherwise.
To characterize the matching constraint, we define a tensor for each $N(v)=\{n_1, n_2,\ldots, n_{d(v)}\}$ as
\begin{equation}
    W_{\langle v, n_1\rangle, \langle v, n_2 \rangle, \ldots, \langle v, n_{d(v)}\rangle} = \begin{cases}
        1, & \sum_{i=1}^{d(v)} \langle v, n_i \rangle \leq 1,\\
        0, & \text{otherwise},
    \end{cases}
\end{equation}
and a tensor for each bond as
\begin{equation}
    B(x^{w_{\langle u,v \rangle}}_{\langle u,v\rangle})_{\langle u, v\rangle} = \begin{cases}
    1, & \langle u, v \rangle = 0 \\
    x^{w_{\langle u,v \rangle}}_{\langle u, v\rangle}, & \langle u, v \rangle = 1,
\end{cases}
\end{equation}
where a label $\langle v, u \rangle$ is equivalent to $\langle u,v\rangle$.
$W$ tensor specifies the constraint that a vertex cannot be shared by two edges in the edge set,
and an edge tensor carries the weights.
Let $x_{\langle u,v\rangle}^{w_{\langle u,v\rangle}}=x$, the tensor network contraction corresponds to the matching polynomial
\begin{equation}
    M(G, x) = \sum\limits_{k=1}^{|V|/2} c_k x^k,
\end{equation}
where $k$ is the size of an edge set, and a coefficients $c_k$ is the number of $k$-matchings.

\subsection{Vertex coloring}
Let $G=(V,E)$ be a graph, a vertex coloring is an assignment of colors to each vertex $v\in V$ such that no edge connects two identically colored vertices. 
In a $k$-coloring problem, the number of colors is limited to less or equal to $k$.
Let us use the 3-coloring problem as an example to show how to reduce it to tensor contractions.
We first map a vertex $v \in V$ to a degree of freedom $c_v\in\{0,1,2\}$.
Then we define a tensor labelled by $c_v$ for each vertex $v$ as
\begin{equation}
    W = \left(\begin{matrix}
        r_v\\
        g_v\\
        b_v
    \end{matrix}\right),
\end{equation}
and a bond tensor labelled by $(c_u, c_v) $ for each edge $(u, v) \in E$ as
\begin{equation}
    B(x^{w_{uv}}) = \left(\begin{matrix}
        0 & x^{w_{uv}} & x^{w_{uv}}\\
        x^{w_{uv}} & 0 & x^{w_{uv}}\\
        x^{w_{uv}} & x^{w_{uv}} & 0
    \end{matrix}\right),
\end{equation}
where $r_v$, $g_v$ and $b_v$ are colors for labeling the configurations. $B$ tensors are for specifying the coloring constraints and $W$ tensors are for labeling the solutions.
%The number of possible colorings can be obtained by contracting this tensor network by setting $r_v, g_v$ and $b_v$ to $1$.
Let $x^{w_{uv}} = x$ and $r_v = g_v = b_v = 1$, we have a graph polynomial,
in which the $k$-th coefficient is the number of coloring that $k$ bonds satisfy constraints.
If a graph is colorable, the maximum order of this polynomial should be equal to the number of edges in this graph.
Similarly, one can define an edge coloring problem by defining the tensor network on the line graph of $G$.

\subsection{Cutting problem}
In graph theory, a cut is a partition of the vertices of a graph into two disjoint subsets,
which is also known as the spin glass problem in statistic physics.
Let $G=(V,E)$ be a graph, we associate a weight $w_v$ to each $v\in V$. To reduce the cutting problem on $G$ to the contraction of a tensor network, we first define a boolean degree of freedom $s_v\in\{0, 1\}$ for each vertex $v\in V$.
Then for each edge $(u,v)\in E$, we define an edge matrix labelled by $(s_u, s_v)$ as
\begin{equation}
    B(x^{w_{uv}}_u, x^{w_{uv}}_v) = \left(\begin{matrix}
        1 & x_{v}^{w_{uv}}\\
        x_{u}^{w_{uv}} & 1
    \end{matrix}\right),
\end{equation}
where variables $x_u^{w_{uv}}$ and $x_v^{w_{uv}}$ are for a cut on this edge or a domain wall in an spin glass.
Let $x_u^{w_{uv}} = x_v^{w_{uv}} = x$, we have a graph polynomial similar to previous ones,
in which the $k$th coefficient is two times the number of cut configurations having size $k$ (i.e. cutting $k$ edges).
